### 2016-08-08 05_conclusions/conclusions.tex
Add the outline and the frontmatter

\section{Conclusions and future works}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../main"
%%% End: