### 2016-08-08 main.tex

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

% Use the acro package to handle the acronym
\usepackage{acro}
\input{00_files_system/acro_def.tex}

\journal{Medical Image Analysis}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{00_files_system/model1-num-names}

%% Numbered without titles
%\bibliographystyle{00_files_system/model1a-num-names}

%% Harvard
\bibliographystyle{00_files_system/model2-names}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{00_files_system/model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{00_files_system/model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{00_files_system/model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{00_files_system/model6-num-names}

%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Automatic prostate cancer detection through \acs*{dce}-\acs*{mri} images: all you need is a good normalization}

% Add authors and the different addresses
\author[label1,label3]{Guillaume~Lema\^itre\corref{cor1}}
\ead{g.lemaitre58@gmail.com}
\author[label3]{Robert~Mart\'i}
%\ead{marly@eia.udg.edu}
\author[label3]{Jordi~Freixenet}
%\ead{jordif@eia.udg.edu}
\author[label5]{Anke Meyer-Baese}
%\ead{ameyerbaese@fsu.edu}
\author[label4]{Joan~C.~Vilanova}
%\ead{kvilanova@comg.cat}
\author[label2]{Paul~M.~Walker}
%\ead{pwalker@u-bourgogne.fr}
\author[label1,label6]{Fabrice~Meriaudeau}
%\ead{fabrice.meriaudeau@u-bourgogne.fr}
\address[label1]{\scriptsize LE2I UMR6306, CNRS, Arts et M\'etiers, Univ. Bourgogne Franche-Comt\'e, 12 rue de la Fonderie, 71200 Le Creusot, France}
\address[label2]{\scriptsize LE2I UMR6306, CNRS, Arts et M\'etiers, Univ. Bourgogne Franche-Comt\'e, Avenue Alain Savary, 21000 Dijon, France}
\address[label3]{\scriptsize ViCOROB, Universitat de Girona, Campus Montilivi, Edifici P4, 17071 Girona, Spain}
\address[label4]{\scriptsize Department of Magnetic Resonance, Cl\'inica Girona, Lorenzana 36, 17002 Girona, Spain}
\address[label5]{\scriptsize Department of Scientific Computing, 400 Dirac Science Library, Florida State University, Tallahassee, FL 32306, United States}
\address[label6]{\scriptsize CISIR, Electrical \& Electronic Engineering Department, Universiti Teknologi Petronas, 32610 Seri Iskandar, Perak, Malaysia}
\cortext[cor1]{Corresponding author.}

% Start the abstract
\begin{abstract}
This template helps you to create a properly formatted \LaTeX\ manuscript.
\end{abstract}

% Give the different keywords
\begin{keyword}
\acs*{dce}-\acs*{mri}\sep prostate cancer\sep normalization\sep classification\sep quantification
\end{keyword}

\end{frontmatter}

% Reset all the acronyms
\acresetall
\linenumbers
% Include now the different files
\input{01_introduction/introduction}
\input{02_methods/methods}
\input{03_experiments/experiments}
\input{04_discussions/discussions}
\input{05_conclusions/conclusions}

% Include the bibliography
\section*{References}

\bibliography{00_files_system/literature}

\end{document}