### 2016-08-09 05_conclusions/conclusions.tex
Finish the introduction

\section{Conclusions and future works}\label{sec:conclusions}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../main"
%%% End: