### 2016-08-22 05_conclusions/conclusions.tex
Add some draft discussion and conclusion

\section{Conclusions and future works}\label{sec:conclusions}

In this work, we presented a new method for normalizing/standardizing \ac{dce}-\ac{mri} data.
Through different experiments, we show the benefit of such normalization.
Furthermore, we show that using the whole \ac{dce}-\ac{mri} signal instead extracting parameters lead to the best classification performance.

As avenues for future research, this normalization has to be part of a \ac{mpmri} \ac{cad} system, mixing \ac{t2w}-\ac{mri}, \ac{dce}-\ac{mri}, \ac{dw}-\ac{mri}, and \ac{mrsi}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../main"
%%% End: