### 2016-09-01 full_title.tex

\documentclass[review, 12pt]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

% Package to change into landscape
\usepackage{pdflscape}

% Package for the appendix
%\usepackage[toc,page]{appendix}
\usepackage[header,title]{appendix}

% Package for table
\usepackage{booktabs}

% For some math expression
\usepackage{amsmath}
\usepackage{mathabx}
\DeclareMathOperator*{\argmin}{arg\,min}

% Package for SI units
\usepackage{siunitx}

% Package for figures
\usepackage{epsf,graphicx}
\usepackage{epstopdf}
\usepackage{subfigure}

% Package for pseudo algorithm
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\makeatletter
\def\BState{\State\hskip-\ALG@thistlm}
\makeatother

% Define the operator of convolution
\newcommand{\Conv}{\mathop{\scalebox{1.5}{\raisebox{-0.2ex}{$\ast$}}}}%

% Use the acro package to handle the acronym
\usepackage{acro}
\input{00_files_system/acro_def.tex}

\journal{Journal of Magnetic Resonance Imaging}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
\bibliographystyle{00_files_system/model1-num-names}

%% Numbered without titles
%\bibliographystyle{00_files_system/model1a-num-names}

%% Harvard
%\bibliographystyle{00_files_system/model2-names}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{00_files_system/numcompress}\bibliographystyle{00_files_system/model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{00_files_system/model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{00_files_system/model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{00_files_system/model6-num-names}

%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Automatic prostate cancer detection through \acs*{dce}-\acs*{mri} images: all you need is a good normalization}

% Add authors and the different addresses
\author[label1,label3]{Guillaume~Lema\^itre\corref{cor1}, MS}
\ead{g.lemaitre58@gmail.com}
\author[label3]{Robert~Mart\'i, PhD}
\author[label1]{Mojdeh~Rastgoo, PhD}
\author[label1]{Joan~Massich, PhD}
%\ead{marly@eia.udg.edu}
\author[label3]{Jordi~Freixenet, PhD}
%\ead{jordif@eia.udg.edu}
\author[label5]{Anke Meyer-Baese, PhD}
%\ead{ameyerbaese@fsu.edu}
\author[label4]{Joan~C.~Vilanova, MD}
%\ead{kvilanova@comg.cat}
%\author[label2]{Paul~M.~Walker}
%\ead{pwalker@u-bourgogne.fr}
\author[label1,label6]{Fabrice~Meriaudeau, PhD}
%\ead{fabrice.meriaudeau@u-bourgogne.fr}
\address[label1]{\scriptsize LE2I UMR6306, CNRS, Arts et M\'etiers, Univ. Bourgogne Franche-Comt\'e, 12 rue de la Fonderie, 71200 Le Creusot, France}
%\address[label2]{\scriptsize LE2I UMR6306, CNRS, Arts et M\'etiers, Univ. Bourgogne Franche-Comt\'e, Avenue Alain Savary, 21000 Dijon, France}
\address[label3]{\scriptsize ViCOROB, Universitat de Girona, Campus Montilivi, Edifici P4, 17071 Girona, Spain}
\address[label4]{\scriptsize Department of Magnetic Resonance, Cl\'inica Girona, Lorenzana 36, 17002 Girona, Spain}
\address[label5]{\scriptsize Department of Scientific Computing, 400 Dirac Science Library, Florida State University, Tallahassee, FL 32306, United States}
\address[label6]{\scriptsize CISIR, Electrical \& Electronic Engineering Department, Universiti Teknologi Petronas, 32610 Seri Iskandar, Perak, Malaysia}
\cortext[cor1]{Corresponding author: Guillaume Lema\^itre, 12 rue de la Fonderie, 71200 Le Creusot, +33761104782}

\begin{abstract}
\begin{description}
  \item[Acknowledgements] None.
  \item[Grant support] Guillaume Lema\^itre is supported by the Generalitat de Catalunya (grant nb. FI-DGR2012) and partly by the Mediterranean Office for Youth (grant nb. 2011/018/06).
  \item[Running title] All you need is a good normalization
\end{description}
\end{abstract}

\end{frontmatter}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: