### 2017-01-27 03_experiments/experiments.tex

\section{Materials}\label{sec:materials}

\subsection{Data}\label{sec:data}

The multi-parametric \ac{mri} data are acquired from a cohort of patients with higher-than-normal levels of \ac{psa}.
Acquisition is achieved with a 3T whole body \ac{mri} scanner (Siemens Magnetom Trio TIM, Erlangen, Germany) using sequences to obtain \ac{t2w}-\ac{mri}, \ac{dce}-\ac{mri} and \ac{dw}-\ac{mri}.
In addition to the \ac{mri} examination, these patients also have
undergone a \ac{trus} guided-biopsy.
The dataset is composed of a total of 20 patients, 18 of which have
biopsies that were positive for \ac{cap} and 2 patients are considered
``healthy'' because they have negative biopsies.
Therefore, 13 patients have \ac{cap} in the \ac{pz}, 3 patients have
\ac{cap} in the \ac{cg}, 2 patients have invasive \ac{cap} in both the
\ac{pz} and the \ac{cg}, and finally 2 patients are considered ``healthy''.
An experienced radiologist segmented the prostate organ---on
\ac{t2w}-\ac{mri} and \ac{dce}-\ac{mri}---as well as the prostate
zones (i.e., \ac{pz} and \ac{cg}) and the \ac{cap} on the \ac{t2w}-\ac{mri}.

A \SI{3}{\mm} slice of fat-suppressed \ac{t2w} fast spin-echo sequence
(\ac{tr}/\ac{te}/\ac{etl}: \SI{3400}{\ms}/\SI{85}{\ms}/13) is used to
acquire images in the sagittal and oblique coronal planes, with the latter planes orientated perpendicular or parallel to the prostate \ac{pz} – rectal wall axis.
Three-dimensional \ac{t2w} fast spin-echo (\ac{tr}/\ac{te}/\ac{etl}: \SI{3600}{\ms}/\SI{143}{\ms}/109, slice thickness: \SI{1.25}{\mm}) images are then acquired in an oblique axial plane.
The nominal matrix and the \ac{fov} of the 3D \ac{t2w} fast spin-echo images are $320 \times 256$ and $280 \times 240$ mm\textsuperscript{2}, respectively, thereby affording sub-millimetric pixel resolution within the imaging plane.

\ac{dce}-\ac{mri} is performed using a fat-suppressed 3D T$_1$ VIBE
sequence (\ac{tr}/\ac{te}/Flip angle:
\SI{3.25}{\ms}/\SI{1.12}{\ms}/\SI{10}{\degree}; Matrix: $256 \times
192$; \ac{fov}: $280 \times 210$ (with 75\% rectangular \ac{fov}); a slab of 16 partitions of \SI{3.5}{\mm} thickness; temporal resolution: \SI{6}{\s}/slab over approximately \SI{5}{\minute}).
A power injector (Medrad, Indianola, USA) is used to provide a bolus injection of Gd-DTPA (Dotarem, Guerbet, Roissy, France) at a dose of \SI{0.2}{\ml} Gd-DTPA/kg of body weight.

These \ac{dce}-\ac{mri} sequences are resampled using the spatial
information of the \ac{t2w}-\ac{mri} and any missing data are added using linear interpolation.
The volumes of the \ac{dce}-\ac{mri} dynamic are rigidly registered to remove any patient motion during the acquisition.
Furthermore, a non-rigid registration is performed between the \ac{t2w}-\ac{mri} and \ac{dce}-\ac{mri} to propagate the prostate zones and \ac{cap} ground-truths.
The resampling is implemented in C++ using the Insight Segmentation and Registration Toolkit~\citep{ibanez2005itk}.

\subsection{Implementation}

The implementation of the registration (C++), normalization (Python), and classification pipeline (Python) are publicly available on GitHub\footnote{\url{https://github.com/I2Cvb/lemaitre-2016-nov/tree/master}}~\citep{lemaitre2016github}.
The data used in this work are also publicly available\footnote{\url{https://zenodo.org/record/61163}}~\citep{lemaitre2016dce}.

\section{Experiments and results}\label{sec:experiments}

\subsection{Goodness of model fitting}\label{sec:fit}

%{\color{red} In case that we have issue with $R^2$, we need to provide the AIC since that the model are usually non-linear.}

\begin{table*}
  \caption{Coefficient of determination $R^{2}$ (i.e., $\mu \ (\pm \sigma)$), while fitting data with the different quantification models.}
  \centering
  \resizebox{\columnwidth}{!}{
  \begin{tabular}{lc c c c c c}
    \toprule
    Data type & Brix & Hoffmann & Tofts population \ac{aif} & Tofts patient \ac{aif} & \ac{pun} & Huisman \\
    \midrule
    Un-normalized & $0.85 \ (\pm 0.11)$ & $0.81 \ (\pm 0.17)$ & $0.84 \ (\pm 0.14)$ & $0.88 \ (\pm 0.12)$ & $0.27 \ (\pm 0.18)$ & $0.64 \ (\pm 0.24)$  \\
    Normalized    & $0.92 \ (\pm 0.05)$ & $0.72 \ (\pm 0.32)$ & $0.92 \ (\pm 0.06)$ & $0.90 \ (\pm 0.10)$ & $0.28 \ (\pm 0.20)$ & $0.75 \ (\pm 0.20)$  \\
    \bottomrule
  \end{tabular}
  }
  \label{tab:r2}
\end{table*}

The parameter estimations from the quantification methods are related to fit a specific model to the \ac{dce}-\ac{mri} data.
Therefore, the coefficient of determination $R^2$ reports the goodness
of fit, as follows:

\begin{equation}
  R^2 = 1 - \frac{\sum_{t = 1}^{T} (s_t - \hat{s}_t)^2}{\sum_{t = 1}^{T} (s_t - \bar{s})^2} ,
  \label{eq:r2}
\end{equation}

\noindent where $s_t$ and $\hat{s}_t$ are the signal to be fitted and
the estimated signal at time $t$, respectively, and $\bar{s}$ is the average signal to be fitted.

The mean and the standard-deviation of the coefficient of determination, $R^{2}$, are reported in Table~\ref{tab:r2} for each quantification model.
Brix, Hoffmann, and Tofts models are fitted with a coefficient $R^{2}$ superior to 0.80.
Additionally, the proposed \ac{pun} model does not to fit the data well.
After introspection of the fitted curves, the original model --- as
formulated in Eq.\,\eqref{eq:pun} --- does not provide enough degrees
of freedom to fit well the data. Additional parameters should be
integrated in this model to control the translation and amplitude of
the model to be fitted.
Data normalization improves the coefficient $R^2$ for all the methods
except for the Hoffmann model.
The large standard deviation for this model might imply that there are some cases where the fitting fails.

% \subsection{Detection of \acs*{cap} using pharmacokinetic parameters}

\subsection{\acs*{cap} detection using pharcokinetic,
  semi-quantitative, and entire enhanced signal}

\begin{table*}
  \caption{\acs*{auc} (i.e., $\mu \ (\pm \sigma)$) for each individual pharmacokinetic parameter using a \acs*{rf} classifier.}
  \centering
  \resizebox{\columnwidth}{!}{
  \begin{tabular}{l c c}
    \toprule
    \textbf{Features} & Un-normalized data & Normalized data \\
    \midrule
    \textbf{Brix model} & & \\
    \quad $A$         & $0.540\ (\pm 0.069)$ & $0.555\ (\pm 0.080)$ \\
    \quad $k_{el}$    & $0.549\ (\pm 0.062)$ & $0.577\ (\pm 0.093)$ \\
    \quad $k_{ep}$    & $0.506\ (\pm 0.032)$ & $0.497\ (\pm 0.019)$ \\
    \textbf{Hoffmann model} & & \\
    \quad $A$         & $0.516\ (\pm 0.020)$ & $0.508\ (\pm 0.031)$ \\
    \quad $k_{el}$    & $0.545\ (\pm 0.066)$ & $0.529\ (\pm 0.065)$ \\
    \quad $k_{ep}$    & $0.550\ (\pm 0.063)$ & $0.545\ (\pm 0.060)$ \\
    \textbf{Tofts model with population \ac{aif}} & & \\
    \quad $K_{trans}$ & $0.556\ (\pm 0.086)$ & $0.565\ (\pm 0.097)$ \\
    \quad $k_{ep}$    & $0.506\ (\pm 0.026)$ & $0.528\ (\pm 0.038)$ \\
    \quad $v_{p}$     & $0.533\ (\pm 0.064)$ & $0.548\ (\pm 0.082)$ \\
    \textbf{Tofts model with patient \ac{aif}} & & \\
    \quad $K_{trans}$ & $0.563\ (\pm 0.077)$ & $0.548\ (\pm 0.060)$ \\
    \quad $k_{ep}$    & $0.492\ (\pm 0.025)$ & $0.491\ (\pm 0.020)$ \\
    \quad $v_{p}$     & $0.530\ (\pm 0.069)$ & $0.495\ (\pm 0.033)$ \\
    \textbf{\ac{pun} model} & & \\
    \quad $a_0$       & $0.521\ (\pm 0.040)$ & $0.530\ (\pm 0.045)$ \\
    \quad $r$         & $0.550\ (\pm 0.085)$ & $0.573\ (\pm 0.097)$ \\
    \quad $\beta$     & $0.531\ (\pm 0.051)$ & $0.549\ (\pm 0.068)$ \\
    \textbf{Semi-quantitative analysis} & & \\
    \quad wash-in     & $0.587\ (\pm 0.107)$ & $0.533\ (\pm 0.032)$ \\
    \quad wash-out    & $0.516\ (\pm 0.037)$ & $0.486\ (\pm 0.035)$ \\
    \quad IAUC        & $0.506\ (\pm 0.048)$ & $0.513\ (\pm 0.032)$ \\
    \quad $\tau$      & $0.565\ (\pm 0.104)$ & $0.537\ (\pm 0.089)$ \\
    \quad $S_M - S_0$ & $0.560\ (\pm 0.083)$ & $0.532\ (\pm 0.029)$ \\
    \bottomrule
  \end{tabular}
  }
  \label{tab:resfeats}
\end{table*}

\begin{figure*}
  \centering
  \hspace*{\fill}
  \subfigure[Quantiative model without normalization.]{\label{fig:rfpharmaunorm}\includegraphics[width=.49\textwidth]{03_experiments/figures/unormalized/unormalized_methods_0.pdf}} \hfill
  \subfigure[Quantitative model With normalization.]{\label{fig:rfpharmanorm}\includegraphics[width=.49\textwidth]{03_experiments/figures/normalized/normalized_methods_0.pdf}}
  \hspace*{\fill}
  \hspace*{\fill}
  \subfigure[Entire enhanced \acs*{dce} signal classification.]{\label{fig:rfnormdcesignal}\includegraphics[width=.49\textwidth]{03_experiments/figures/full_signal_0.pdf}}
  \hspace*{\fill}
  \caption{\acs*{roc} analysis using a \acs*{rf} classifier with and
    without normalization using different approaches: \protect\subref{fig:rfpharmaunorm} -
    \protect\subref{fig:rfpharmanorm} pharmacokinetic and
    semi-quantitative models without and with normalization,
    respectively; \protect\subref{fig:rfnormdcesignal} entire enhanced
  \acs*{dce}-\acs*{mri} signal.}
  \label{fig:normpharmarf}
\end{figure*}

\begin{figure}
  \centering
  \includegraphics[width=1.\linewidth]{03_experiments/figures/wilcoxon.pdf}
  \caption{Wilcoxon signed-ranked test to compare each pair of
    classifiers. The annotation in the matrix corresponds to the
    p-values ($p$). $p < 0.05$ indicates that a pair of classifiers
    are significantly different, consequently the classifier with the
    highest \acs*{auc} significantly outperforms the other
    one. In the co-occurrence matrix, {\color{blue}blue} cells
    correspond to lower $p$ while {\color{red} red} cells are
    representing greater $p$.}
  \label{fig:wilcoxon}
\end{figure}

To study the potential benefit of our normalization, \ac{cap} is detected at a voxel level using pharmacokinetic parameters estimated from non-normalized and normalized \ac{dce}-\ac{mri} data.
Each individual pharmacokinetic parameter is classified to evaluate its individual discriminative power to detect \ac{cap}.
In addition, p-values ($p$) are computed to identify pairs of classifiers
which are significantly different, using a Wilcoxon signed-ranked
test. Those p-values are represented in the color coded co-occurrence
matrix shown in Fig.\,\ref{fig:wilcoxon}, where blue cells correspond
to lower p-values while red cells are greater p-values.
Therefore, a \ac{rf} classifier is used in conjunction with a \ac{lopo}.
The use of \ac{rf} is encouraged since it leads to the best performance in the state-of-the-art methods~\citep{litjens2014computer,lemaitre2015computer}.
The results are summarized in Table~\ref{tab:resfeats} in terms of the \ac{auc}.
Normalization can improve the detection of \ac{cap}, but the benefit
of normalization is more obvious with the combination of the
pharmacokinetic features of a given model (e.g., $A$, $k_ep$, and
$k_el$ for the Brix model) as previously executed in the traditional \ac{cad} system~\citep{lemaitre2015computer}.
For the latter configuration, results are summarized by performing a \ac{roc} analysis and computing the \ac{auc}, as reported in Fig.\,\ref{fig:normpharmarf}.
Quantification using normalized data outperforms quantification using
non-normalized data in terms of the classification performance, except
for Hoffmann and Tofts population-based \ac{aif} models.
The reasons behind the decrease of the \ac{auc} might be related to:
(i) a poor fitting as discussed in Sect.\,\ref{sec:fit} (cf., Hoffmann
model) and (ii) a small number of patients for the estimation of some parameters (cf., Tofts model).
The best classification performance is obtained using the
semi-quantitative approach with an \ac{auc} of $0.655 \pm 0.106$.

%\subsection{Classification of the entire enhanced \acs*{dce}-\acs*{mri} signal}

\begin{figure*}
  \centering
  \includegraphics[width=1.\linewidth]{03_experiments/figures/plot_auc_by_zones.pdf}
  \caption{Comparison of the distribution of the \acs*{auc} score for
    each prostate zone, with and without normalization.}
  \label{fig:boxplot}
\end{figure*}

As stated in the introduction, the quantification methods extract a
set of parameters that characterize the enhancement of the \ac{dce}-\ac{mri} signal.
However, this extraction might lead to a loss of information.
This experiment is performed to assess whether using the whole
\ac{dce}-\ac{mri} signal, rather than just the pharmacokinetic parameters, can improve the classification performance.
Therefore, each enhanced \ac{dce}-\ac{mri} signal, normalized and un-normalized, is classified using a \ac{rf} classifier in a \ac{lopo} fashion.
The \ac{roc} analysis and \ac{auc} are reported in Fig.\,\ref{fig:rfnormdcesignal}.
Classification without normalization leads to the worst performance,
with an \ac{auc} of $0.568 \pm 0.162$.
However, data normalization in conjunction with the use of the entire
\ac{dce}-\ac{mri} signal is the strategy that outperforms others,
with an \ac{auc} of $0.666 \pm 0.154$.
Comparable outcomes are visible while analyzing the \ac{auc}
scores distribution for each prostate zone as shown in
Fig.\,\ref{fig:boxplot}. The \ac{auc} scores
related to \ac{cap} in \ac{cg} improves using normalized data, for
most of the models.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: