### 2017-01-27 05_conclusions/conclusions.tex

\section{Conclusions and future works}\label{sec:conclusions}

In this work, we presented a new method for normalizing/standardizing \ac{dce}-\ac{mri} data.
This method is designed to reduce the inter-patient variations that occur during data acquisition.
A graph-based approach was used to correct intensity offset along with a model-based correction to reduce time offset and intensity scaling.
We show the benefit of our normalization method prior to extract quantitative and semi-quantitative features, with a significant improvement of the classification performance.
Nevertheless, we also show that using the whole normalized \ac{dce}-\ac{mri} signal outperforms all quantitative approaches.

In future research, this normalization needs to be part of an
\ac{mpmri} \ac{cad} system in which the \ac{dce}-\ac{mri} modality needs to be combined with other complementary modalities.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: