### 2017-01-31 response_to_reviewers.tex
add response to reviewer

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Thin Formal Letter
% LaTeX Template
% Version 1.11 (8/12/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% WikiBooks (http://en.wikibooks.org/wiki/LaTeX/Letters)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% ----------------------------------------------------------------------------------------
% DOCUMENT CONFIGURATIONS
% ----------------------------------------------------------------------------------------

\documentclass{letter}

% Adjust margins for aesthetics
\addtolength{\voffset}{-0.5in}
%\addtolength{\hoffset}{-0.3in}
\addtolength{\textheight}{2cm}

\usepackage[letterpaper, margin=1.5in]{geometry}

%\longindentation=0pt % Un-commenting this line will push the closing "Sincerely," to the left of the page

% ----------------------------------------------------------------------------------------
% YOUR NAME & ADDRESS SECTION
% ----------------------------------------------------------------------------------------

\signature{Guillaume Lema\^itre} % Your name for the signature at the bottom

\address{Parietal team, Inria \\ CEA, Universit\'e Paris-Saclay \\ 1 Rue
  Honor\'e d'Estienne d'Orves \\ 91120 Palaiseau \\ France} % Your address
% and phone number

% ----------------------------------------------------------------------------------------

\begin{document}

% ----------------------------------------------------------------------------------------
% ADDRESSEE SECTION
% ----------------------------------------------------------------------------------------

\begin{letter}{Nicholas Ayache \\ Co-editor in Chief \\ Medical Image Analysis} % Name/title of the addressee

  % ----------------------------------------------------------------------------------------
  % LETTER CONTENT SECTION
  % ----------------------------------------------------------------------------------------

  \opening{
    \textit{Subject: MEDIA-D-16-00507 - Response to reviewers}\\ \\
    \textbf{Dear Sir or Madam,}}

  We would like to thank the reviewers for their pertinent
  remarks that have increased the quality of the manuscript. Whenever
  possible, the comments of the reviewers have been addressed and
  highlighted in the article.

  Please, find below a point-by-point response to each issues raised
  by the reviewers.

  Reviewer \#1:
  \begin{enumerate}
  \item \textbf{In a typical DCE based analysis, the contrast
      enhanced images are rigidly registered to T1w images for
      intensity normalization. Typically for multi-parametric MRI
      based CAD system for prostate cancer detection T1w images
      are present. This is useful in differentiating cancers from
      blood hemorrhage following prostate biopsy. Please discuss
      why a pre or a post contrast T1w image is missing from the
      clinical trial that will be using DCE images.}\\
    The MRI acquisition was performed prior to carry out the
    biopsy. Therefore, the risk of blood hemorrhage is discarded.
    Additionally, pre-contrast T1w images are available since 4 to 6
    volumes are acquired previous to the bolus injection of the
    contrast media. Indeed, this pre-contrast data are used in some
    models (i.e., $S_0$).
  \item \textbf{The other advantage of normalizing to T1w images are
      patient specific information will be preserved. Different
      patients may have different cancer lesions of different grades
      (having different kinetic expression in DCE) and normalizing all
      patients to the same space may result in loss of
      information. Please comment.}\\
    Our normalization is designed to align the \emph{mean kinetic
      expression}. Therefore, it is true that patient-related
    information can be lost. However, lesion with different grades
    should more or less dispersed to the mean kinetic expression. This
    information remains available after applying our normalization.
  \item \textbf{T1w normalization will be probably be more robust to
      change in magnet, machine, reconstruction software and
      acquisition parameters across different institutions as the
      objective is normalization to be patient and institution
      specific. Please comment how the presented model may be affected
      in a multi-institutional study.}\\
    Our method does not rely on any manufacturer/machine-specific settings.
    Subsequently, our approach should work on any type
    of devices, enabling multi-institutional studies. We included
    these remarks in our future work section in p.~23/l.~396.
  \item \textbf{Please discuss how the improvement in prostate cancer
      detection in DCE-MRI impacts the current clinical CAD
      pipeline. According to PIRADS v2 guidelines more emphasis is
      laid upon T1w, T2w and DWI compared to DCE images. Experienced
      radiologists can detect prostate cancers in peripheral zone with
      a high degree of accuracy (higher than 0.6 AUC). The transition
      zone lesions are more difficult to distinguish. Please comment
      how the model performs in detection of transition zone
      lesions. Further please comment how this model adds value to
      already established clinical pipeline of detecting prostate
      cancer (PIRADS).}\\
    A stratified analysis separating PZ and CG showed an increase of
    the classification performance in CG after applying our
    normalization method as depicted in Fig.\,8 p.~22.
    The PIRADS v2 emphasized that DCE-MRI have to be included in
    mp-MRI acquisition to detect small lesion. However, it also warn
    that the benefit is modest. Therefore, the impact of an
    improvement of detection in DCE-MRI can only be evaluated in a
    fully mp-MRI CAD system. Indeed, we recently showed, --- Lemaitre,
    Guillaume. Computer-Aided Diagnosis for Prostate Cancer using
    Multi-Parametric Magnetic Resonance Imaging. Diss. Universit\'e de
    Bourgogne; Universitat de Girona, 2016;
    Chapt.\,6/p.~150/Table~6.10 ---, that DCE-MRI features are
    considered important and the combination with other modalities
    will lead to better classification results.
    However, in our humble opinion, including those additional results
    will clutter the paper.
  \item \textbf{Comparing the results of semi-automatic approach and
      the presented approach of curve fitting using the entire curve
      the results are not significantly different (0.65 AUC vs 0.66
      AUC). The presented model is however automatic which would
      reduce inter-rater variance in a semi-automatic method. Please
      present a discussion comparing these two results.}\\
    It is important to notice that for both approaches, ---
    semi-quantitative and entire enhanced signal ---, the results
    are obtained using our normalization approach, showing the importance of
    this step. We added a discussion in p.~23/l.~388 regarding the
    semi-quantitative approach and entire enhanced signal
    classification. The semi-quantitative approach reduce the number
    of features to analyze, speeding up the classification. However,
    it relies on the curve fitting which is also time consuming and
    might fail.
  \item \textbf{Use of single feature for a random forest classifier
      (Table 2) may be detrimental for the classifier as random
      sub-sampling of the features are not possible. Please present
      the comparison with some linear classifiers like linear SVM.}\\
    A classification using a linear SVM has been added in additional
    material.
    Unlike a linear SVM, RF allows to find multiple thresholds,
    leading to a much complex decision boundary for this
    feature. Bootstrapping the samples for each tree allows to cope
    with the over-fitting of each decision trees. Therefore, we do not
    think that a linear SVM will substantially better insights as are
    showing the results provided in the additional material.
  \end{enumerate}

  Reviewer \#2:
  \begin{enumerate}
  \item \textbf{The title is misleading and not very scientific and I
      would recommend to remove the second part and reword the first
      part to cover the contents of the paper.}\\
    We are not against changing the title but we would like to take
    advantage of this revision to get the opinion of all reviewers
    regarding this matter.
  \item \textbf{The use of the English language needs some attention:
      get it all proof read.}\\
    As suggested, the manuscript has been corrected by an English editing
    service to ensure a good quality writing.
  \item \textbf{What guided biopsy was used: US?}\\
    The guided biopsy was performed with a TRUS. We added this
    statement in p.~15/l.~275.
  \item \textbf{The number of cases seems low. This is mentioned as
      possibly affecting the results and that the limitations might be
      based on the sample size. At the end of section 5 one could draw
      the conclusion that they should do a more extensive study.}\\
    The statement has been added in the discussions p.~23/l.~396.
  \item \textbf{How would this translate to other types of MRI
      systems: it would be good to see experimental results on this or
      at least some discussion.}\\
    Refer to reviewer \#1 / comment 3 since this is a shared remarks.
  \item \textbf{The disappointing performance and large standard
      deviation of the PUN model should be investigated and
      explained.}\\
    A discussion is added in p.~17/l.~320. The original model lacks of
    degrees of freedom. Two additional parameters will
    allow for a better fit. However, we used the original formulation
    in our experiments.
  \item \textbf{There should be a detailed discussion on the results
      to indicate which differences are statistically
      significant. ``Significant'' is mentioned in the conclusions, but
      there is no evidence for this.}\\
    We added a Wilcoxon significance test - see Fig.\,7. Now, the term
    significant is only used for $p<0.05$.
  \item \textbf{The overall results seem poor and this could be
      further discussed.}\\
    It is difficult to compare with the state-of-the-art methods since
    the dataset used are different and the reported results in those
    studies are usually linked to a mp-MRI CAD instead of only
    DCE. Nevertheless, our results are in the same range than Litjens,
    Geert, et al. "Computer-aided detection of prostate cancer in
    MRI." IEEE transactions on medical imaging 33.5 (2014):
    1083-1092 regarding the classification performance of the Tofts
    parameters.
  \item \textbf{Sections 5 and 6 are both disappointingly short.}
    Due to additional analyses, the discussion became longer.
    % However,
    % we believed that there is no need for repeating information
    % already presented in the experiment section or give hypothesis
    % which cannot be verified.
  \item \textbf{Some journals references incomplete. Gliozzi et al. is
      not a single page. Check: Schabel, Shanbhag, Zhu.}\\
    The references have been corrected.
  \end{enumerate}

  Reviewer \#5:
  \begin{enumerate}
  \item \textbf{There are however several issues with the paper In the
      current state. Using quantitative parameters obtained from
      pharmacokinetic models as an end-point comparison indeed may be
      important as the models provide physiological information
      related to capillary permeability, blood fraction and
      interstitial space volume, etc.. However I am not really
      convinced by the detailed results presented in Table 2, as no
      statistical test was performed when the classification
      performances are higher with normalized data, and for some of
      the models the results are worst. For me, the benefits of the
      method are not fully demonstrated in terms of classification
      performance using the models. How these parameters are modified
      with the normalization?}\\
    This is important to emphasize that our goal was to propose a
    normalization method to allow classification of the entire
    enhanced signal which should perform as good as the current
    models. We apply the same normalization on the different
    pharmacokinetic models and, as mentioned by the reviewer ---
    the results are not significantly better.
    Indeed, these experiments were carried out to investigate the
    benefit or not of the normalization with the current models,
    without expecting the results to improve.
  % \item \textbf{Although quantitative parameters obtained from
  %     pharmacokinetic models applied to DCE-MRI series have
  %     demonstrated in general superior performance to differentiate
  %     cancerous from normal tissue prostate in comparison to simpler
  %     enhancement analysis such as semi-quantitative or qualitative
  %     parameters, there are still several differences in results
  %     across the studies in the literature. There are some papers
  %     aimed at defining normal values for some of the pharmacokinetic
  %     parameters. The authors may refer to the paper Sanz-Requena et
  %     al. Dynamic contrast-enhanced case-control analysis in
  %     3T MRI of prostate cancer can help to characterize tumor
  %     aggressiveness, European Journal of Radiology, 85(11), November 2016,
  %     pp. 2119-2126, for a detailed list of typical values and difficulties
  %     of the models}\\
  \item \textbf{Some of the benefits or improvements may be
      demonstrated before any further classification step. The method
      could be useful when performing population analysis with the raw
      data, previous to defining a model.}\\
    We do not really see which analysis can be performed to evaluate
    the benefits of our methods before the classification step. A study of
    the variances with and without normalization could be carried out but
    an evaluation after classification seems equivalent and a good manner to
    evaluate the benefit of the normalization. In fact, this is the
    results reported in Fig.\,6(c).
  \item \textbf{Figure 1 is not pertinent for the goal of this paper.
      It is not clear how is represented the interindividual variation in
      figure 2?  In Fig 2a the PDFs are concatenated in time to obtain a
      heatmap representation, but in 2 b-c  the differences across
      individuals does not appear.}\\
    Figure 1 is used to illustrate inter-patients variation in T2w-MRI
    and to highlight the need of a different normalization approach
    than the one needed in T2w. We improve the Fig.\,2. Figure~2(a)
    presents only the process to build the
    heatmap. In Fig.\,2(b)-2(c), we added parameters which highlight
    more clearly the variations and added their corresponding values
    in caption.
  \item \textbf{It is not clear the rational of equation 1, as the
      parameters are rather empirical.  There are no other ways to
      smooth the heatmap, for example through a diffusion process,
      solving a PDE equation? Or by registering through the series?}\\
    There is probably a confusion. We do not try to smooth the heatmap
    but we search a smooth estimate. Equation 1 represents the weight
    of the graph and is a standard way to formalize graph-based
    problem in image processing.
  \item \textbf{The theoretical foundations of equation 3 are not
      clear. Other registration methods for time signas may have been
      included.}\\
    While other registrations could have been used,
    using a rigid registration directly affect the variations
    $\alpha_i$ and $\Delta_t$. Applying more advanced registration
    such as elastic registration as in Lemaitre, Guillaume, et
    al. "Normalization of T2W-MRI prostate images using Rician a
    priori." SPIE Medical Imaging. International Society for Optics
    and Photonics, 2016. would make it complicated to shift the data
    with only these two parameters. Furthermore, rigid registration
    will not change the variations of the contrast agent over time.
  \item \textbf{Other question that may be raised is the influence of
      the equation parameters (tau, alfa..) in the pharmacokinetic
      models. Do they change with f(t)?}\\
    We think that this question is related to the results presented in
    Sect.\,4.1
  \item \textbf{A thorough review of the structure and in general an
      improvement of the paper in terms of writing may be helpful to
      fully understand the message, the rational and the proposed
      methodology.}\\
    As suggested, the manuscript has been corrected by an English editing
    service to ensure a good quality writing.
  \end{enumerate}

  Thank you for your time and consideration.

  I look forward to your reply.

  \vspace{2\parskip} % Extra whitespace for aesthetics
  \closing{Sincerely,}
  \vspace{2\parskip} % Extra whitespace for aesthetics

  % ----------------------------------------------------------------------------------------

\end{letter}

\end{document}